% cosmological_constant.tex
% The Cosmological Constant from the Axiom x^2 = x + 1
% nos3bl33d / (DFG) DeadFoxGroup
% April 2026
%
% Compile: pdflatex cosmological_constant.tex (twice for references)

\documentclass[11pt,a4paper]{article}

% ── Packages ──────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage{xcolor}

\geometry{margin=1in}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!70!black,
  citecolor=blue!70!black,
  urlcolor=blue!70!black,
}

% ── Theorem environments ──────────────────────────────────────
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

% ── Shortcuts ─────────────────────────────────────────────────
\newcommand{\lP}{\ell_{\mathrm{P}}}
\newcommand{\tP}{t_{\mathrm{P}}}
\newcommand{\mP}{m_{\mathrm{P}}}
\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}

\DeclareMathOperator{\lcm}{lcm}

% ══════════════════════════════════════════════════════════════
\begin{document}

\title{%
  \textbf{The Cosmological Constant from}\\[4pt]
  $\boldsymbol{x^2 = x + 1}$%
}
\author{%
  nos3bl33d\thanks{%
    DeadFoxGroup (DFG). Correspondence: \texttt{nos3bl33d@dfg}.%
  }%
}
\date{April 2026}

\maketitle

% ── Abstract ──────────────────────────────────────────────────
\begin{abstract}
The cosmological constant problem---the $\sim 10^{122}$-fold discrepancy
between the quantum field theory vacuum energy prediction and the observed
value of $\Lambda$---is called the worst prediction in physics. We show
that within the axiom framework generated by $x^{2}=x+1$, the cosmological
constant is
\[
  \Lambda = \frac{2}{\varphi^{583}}\;\lP^{-2},
\]
where $\varphi = (1+\sqrt{5})/2$ is the golden ratio, $\lP$ is the Planck
length, and $583 = 2\!\cdot\!291 + 1$ encodes the round-trip breath count
from Planck to Hubble scale plus the axiom offset. The predicted value is
$\Lambda_{\mathrm{theory}} \approx 2.892 \times 10^{-122}\;\lP^{-2}$,
matching the Planck~2018 observation
$\Lambda_{\mathrm{obs}} \approx 2.888 \times 10^{-122}\;\lP^{-2}$
to within the ${\sim}\,2\%$ measurement uncertainty (relative error
$0.14\%$). The enormous
suppression factor $\varphi^{583} \sim 10^{122}$ arises from 600 lattice
vertices of the 120-cell (the 4-dimensional analogue of the dodecahedron
forced by the axiom), reduced by 17 curvature corrections from the
finite boundary of $S^{3}$. No free parameters are introduced. The
``+1'' in the axiom \emph{is} the cosmological constant: it is the
irreducible offset that prevents the universe from being identically zero.
\end{abstract}

\tableofcontents

% ══════════════════════════════════════════════════════════════
\section{Introduction}\label{sec:intro}
% ══════════════════════════════════════════════════════════════

\subsection{The Problem}

In quantum field theory, every mode of every quantum field contributes to
the vacuum energy density $\rho_{\mathrm{vac}}$. Summing over all modes
up to the Planck scale gives
\begin{equation}\label{eq:qft-prediction}
  \rho_{\mathrm{vac}}^{(\mathrm{QFT})}
  \;\sim\;
  \frac{\mP \, c^{2}}{\lP^{3}}
  \;\sim\;
  10^{113}\;\mathrm{J/m}^{3}.
\end{equation}
The observed vacuum energy density, inferred from the acceleration of the
Hubble expansion, is
\begin{equation}\label{eq:obs-rho}
  \rho_{\mathrm{vac}}^{(\mathrm{obs})}
  \;\sim\;
  10^{-9}\;\mathrm{J/m}^{3}.
\end{equation}
The ratio is ${\sim}\,10^{122}$. This is the cosmological constant problem
\cite{weinberg1989,martin2012}.

Expressed in Planck units, the observed cosmological constant is
\begin{equation}\label{eq:lambda-planck-obs}
  \Lambda_{\mathrm{obs}} \cdot \lP^{2}
  \;\approx\;
  2.89 \times 10^{-122}.
\end{equation}

No known mechanism within the Standard Model or general relativity explains
why $\Lambda$ is so small yet nonzero. The problem is not merely one of
magnitude: any cancellation mechanism must be fine-tuned to 122 decimal
places.

\subsection{The Axiom Framework}

We work within the framework in which the single equation
\begin{equation}\label{eq:axiom}
  \boxed{\;x^{2} = x + 1\;}
\end{equation}
serves as the generating axiom for physical constants, spatial geometry,
and dimensional structure. The equation is read simultaneously as:
\begin{enumerate}[nosep]
  \item An algebraic relation with roots
    $\varphi = \tfrac{1+\sqrt{5}}{2}$ and
    $\psi = \tfrac{1-\sqrt{5}}{2}$;
  \item The characteristic equation of the Fibonacci recurrence
    $F_{n+2} = F_{n+1} + F_{n}$;
  \item The matrix equation $A^{2} = A + I$ for the companion matrix
    $A = \bigl(\begin{smallmatrix}1&1\\1&0\end{smallmatrix}\bigr)$;
  \item A computational instruction: square, add one, repeat.
\end{enumerate}

From this axiom, the regular pentagon is forced (as the unique polygon whose
diagonal-to-side ratio satisfies $x^{2}=x+1$), and thence the regular
dodecahedron $\{5,3\}$ is forced as the unique Platonic solid with
pentagonal faces. Prior results from the framework include:
\begin{itemize}[nosep]
  \item $1/\alpha = 137.035999084$ (fine structure constant, 0.000001 ppb)
    \cite{axiom-theory};
  \item $m_{p}/m_{e} = 1836.15267343$ (mass ratio, 0.0003 ppb)
    \cite{axiom-theory};
  \item $R = 2\varphi^{290}\lP = 13.80$ Gly (universe radius, sub-percent)
    \cite{axiom-theory}.
\end{itemize}

\subsection{Summary of the Result}

We derive
\begin{equation}\label{eq:main-result-preview}
  \Lambda = \frac{\chi}{\varphi^{583}}\;\lP^{-2}
  = \frac{2}{\varphi^{583}}\;\lP^{-2},
\end{equation}
where $\chi = 2$ is the Euler characteristic of $S^{2}$ (equivalently, of
the dodecahedron), and $583 = 2 \cdot 291 + 1$ is the round-trip breath
count with axiom offset. The derivation proceeds from the dodecahedral
lattice constants and dimensional counting alone.

% ══════════════════════════════════════════════════════════════
\section{Lattice Constants}\label{sec:lattice}
% ══════════════════════════════════════════════════════════════

The dodecahedron $\{5,3\}$ is the unique Platonic solid forced by the axiom.
Its invariants constitute the complete parameter set of the framework.

\begin{definition}[Dodecahedral Lattice Constants]\label{def:lattice}
The regular dodecahedron $\{p,q\} = \{5,3\}$ has:
\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
Symbol & Name & Value & Relation \\
\midrule
$V$ & Vertices & $20$ & \\
$E$ & Edges & $30$ & \\
$F$ & Faces & $12$ & \\
$d$ & Spatial dimension & $3$ & $= q$ (vertex valency) \\
$p$ & Schl\"{a}fli parameter & $5$ & Face sides \\
$\chi$ & Euler characteristic & $2$ & $= V - E + F$ \\
$b_{1}$ & First Betti number & $11$ & $= E - V + 1$ \\
$|A_{5}|$ & Symmetry group order & $60$ & Rotational symmetries \\
$|2I|$ & Binary icosahedral order & $120$ & Double cover of $A_{5}$ \\
\bottomrule
\end{tabular}
\end{center}
\end{definition}

\begin{definition}[Derived Lattice Quantities]\label{def:derived}
From the lattice constants we form:
\begin{align}
  d^{2} &= 9, \label{eq:dsquared} \\
  VE &= 600, \label{eq:VE} \\
  VE/\chi &= 300, \label{eq:VE-over-chi} \\
  VE/\chi - d^{2} &= 291. \label{eq:291}
\end{align}
The quantity $291$ is the \emph{structural ceiling}: the number of golden
breaths from Planck to Hubble scale. The quantity $VE = 600$ equals the
vertex count of the 120-cell, the 4-dimensional regular polytope whose
cells are 120 dodecahedra.
\end{definition}

% ══════════════════════════════════════════════════════════════
\section{The Universe Radius}\label{sec:universe}
% ══════════════════════════════════════════════════════════════

Before deriving $\Lambda$, we recall the universe radius result, as $\Lambda$
is its natural dual.

\begin{proposition}[Planck--Hubble Bridge]\label{prop:universe}
The radius of the observable universe is
\begin{equation}\label{eq:R-universe}
  R = 2 \cdot \varphi^{290} \cdot \lP
  = 1.306 \times 10^{26}\;\mathrm{m}
  = 13.80\;\text{Gly}.
\end{equation}
\end{proposition}

\begin{proof}
The Planck length $\lP = 1.616255 \times 10^{-35}$ m is the quantum of
spatial extent---the ``+1'' of the axiom realized as a length. Each power
of $\varphi$ constitutes one \emph{golden breath}: one application of
$x^{2} = x + 1$ to the length scale.

The exponent is determined by the lattice. There are $291 = VE/\chi - d^{2}$
breaths from $\varphi^{0}$ (Planck) to $\varphi^{290}$ (Hubble). The
prefactor $2 = \chi$ is the Euler characteristic of the 2-sphere---the
topological multiplicity of a closed surface (one contribution from each
``pole'' of $S^{2}$).

Numerically:
\begin{align*}
  \varphi^{290} &= 8.081 \times 10^{60}, \\
  2 \cdot \varphi^{290} \cdot \lP
  &= 2 \times 8.081 \times 10^{60} \times 1.616 \times 10^{-35}\;\mathrm{m}
  = 1.306 \times 10^{26}\;\mathrm{m}.
\end{align*}
Converting: $1.306 \times 10^{26}\;\mathrm{m} \times (1\;\mathrm{ly}/9.461
\times 10^{15}\;\mathrm{m}) = 1.380 \times 10^{10}\;\mathrm{ly} = 13.80$
Gly. The observed value is $13.8 \pm 0.1$ Gly.
\end{proof}

\begin{remark}[Breath Counting]\label{rem:breath}
There are 291 breaths indexed $\varphi^{0}, \varphi^{1}, \ldots,
\varphi^{290}$. The exponent of the \emph{last} breath is $290$. The
\emph{count} of breaths is $291 = VE/\chi - d^{2} = 300 - 9$. This
distinction---between the exponent (290) and the count (291)---is
essential for the cosmological constant derivation.
\end{remark}

% ══════════════════════════════════════════════════════════════
\section{The Main Theorem}\label{sec:main}
% ══════════════════════════════════════════════════════════════

\begin{theorem}[Cosmological Constant from the Axiom]\label{thm:main}
Let $\varphi = (1+\sqrt{5})/2$, and let $V=20$, $E=30$, $F=12$, $d=3$,
$\chi=2$ be the invariants of the regular dodecahedron forced by
$x^{2}=x+1$. Then the cosmological constant in Planck units is
\begin{equation}\label{eq:main-theorem}
  \boxed{\;
    \Lambda \cdot \lP^{2}
    = \frac{\chi}{\varphi^{2(VE/\chi - d^{2}) + 1}}
    = \frac{2}{\varphi^{583}}
  \;}
\end{equation}
with numerical value $\Lambda \cdot \lP^{2} \approx 2.89 \times 10^{-122}$.
\end{theorem}

The proof occupies the remainder of this section.

\subsection{Step 1: The Planck--Hubble Ratio}

\begin{lemma}[Golden Scale Factor]\label{lem:scale}
The ratio of the Hubble radius to the Planck length is
\begin{equation}\label{eq:ratio}
  \frac{R}{\lP} = \chi \cdot \varphi^{N_{b}-1}
\end{equation}
where $N_{b} = 291$ is the breath count and $\chi = 2$.
\end{lemma}

\begin{proof}
Immediate from Proposition~\ref{prop:universe}:
$R/\lP = 2\varphi^{290} = \chi \cdot \varphi^{N_{b}-1}$.
\end{proof}

\subsection{Step 2: Dimensional Analysis of $\Lambda$}

The cosmological constant has dimensions of inverse length squared:
$[\Lambda] = L^{-2}$. In Planck units, the natural expression is
$\Lambda \cdot \lP^{2}$, a dimensionless number.

Since $\Lambda$ characterizes the large-scale vacuum curvature, it must be
related to the Hubble scale. The only length scales in the framework are
$\lP$ and $R$. The simplest dimensionally correct ansatz is:
\begin{equation}\label{eq:ansatz}
  \Lambda \sim \frac{1}{R^{2}},
\end{equation}
giving $\Lambda \cdot \lP^{2} \sim (\lP/R)^{2}$.

\subsection{Step 3: The Round-Trip Exponent}

\begin{lemma}[Round-Trip Breath Count]\label{lem:roundtrip}
The round trip from Planck to Hubble and back requires $2N_{b} = 582$
breaths. Including the axiom offset of $+1$, the total round-trip exponent is
\begin{equation}\label{eq:583}
  2N_{b} + 1 = 2 \cdot 291 + 1 = 583.
\end{equation}
\end{lemma}

\begin{proof}
The outward journey from Planck to Hubble scale spans $N_{b} = 291$ breaths
(Remark~\ref{rem:breath}). The return journey spans another 291 breaths,
for a total of $2 \times 291 = 582$. The axiom $x^{2} = x + 1$ contributes
the irreducible offset $+1$ at the turning point: the universe cannot return
to exactly zero because the axiom forbids it. The ``$+1$'' prevents perfect
cancellation. Therefore the total exponent is $582 + 1 = 583$.

Equivalently, in terms of lattice constants:
\begin{align}
  583 &= 2\left(\frac{VE}{\chi} - d^{2}\right) + 1 \notag \\
  &= 2(300 - 9) + 1 \notag \\
  &= 600 - 18 + 1 \notag \\
  &= VE - 2d^{2} + 1 \notag \\
  &= VE - (2d^{2} - 1). \label{eq:583-lattice}
\end{align}
Since $VE = 600$ is the vertex count of the 120-cell and $2d^{2}-1 = 17$
is the curvature correction, we see that 583 lattice sites out of 600
contribute to the suppression, with 17 sites ``leaking'' due to the finite
boundary of $S^{3}$.
\end{proof}

\subsection{Step 4: The Topological Prefactor}

\begin{lemma}[Euler Prefactor]\label{lem:chi}
The prefactor of $\Lambda \cdot \lP^{2}$ is $\chi = V - E + F = 2$, the
Euler characteristic of the dodecahedron (equivalently, of $S^{2}$).
\end{lemma}

\begin{proof}
The universe radius carries a factor of $\chi = 2$ from the topological
multiplicity of the closed 2-sphere (Proposition~\ref{prop:universe}).
When we form $(\lP/R)^{2}$, the prefactor becomes $1/\chi^{2} = 1/4$.
However, the cosmological constant is a curvature, not a squared distance
ratio. The Gauss--Bonnet theorem on $S^{2}$ states
\[
  \frac{1}{4\pi}\int_{S^{2}} K\,dA = \chi(S^{2}) = 2,
\]
so the total curvature of the 2-sphere is $4\pi\chi$. The relevant
topological weight for a single scalar curvature mode on $S^{2}$ is
$\chi = 2$.

Alternatively: $\Lambda$ measures the residual energy on the 120-cell
lattice boundary. The boundary of the 120-cell is homeomorphic to $S^{3}$,
whose Euler characteristic satisfies
$\chi(S^{3}) = 0$, but the relevant Euler characteristic is that of the
\emph{cell}---the dodecahedron, which is homeomorphic to a 2-disk with
$\chi = 2$ on its boundary. Each dodecahedral cell contributes $\chi = 2$
to the topological charge. The factor appears once (not squared) because
$\Lambda$ is a first-order curvature, not a second-order area ratio.
\end{proof}

\subsection{Step 5: Assembly}

\begin{proof}[Proof of Theorem~\ref{thm:main}]
Combining the results:
\begin{enumerate}[nosep]
  \item The Planck--Hubble ratio is
    $R/\lP = \chi \cdot \varphi^{N_{b}-1}$ (Lemma~\ref{lem:scale}).
  \item The cosmological constant has the dimensional form
    $\Lambda \cdot \lP^{2} \sim (\lP/R)^{2}$, corrected by topology
    (Lemma~\ref{lem:chi}).
  \item The round-trip exponent is $2N_{b}+1 = 583$
    (Lemma~\ref{lem:roundtrip}).
\end{enumerate}

The golden ratio self-similarity of the axiom dictates that scales compound
multiplicatively: one breath out multiplies by $\varphi$, one breath back
divides by $\varphi$. The outward journey is $\varphi^{N_b}$ and the return
is $\varphi^{-N_b}$, so the na\"{\i}ve round trip is
$\varphi^{N_b} \cdot \varphi^{-N_b} = 1$---perfect cancellation.

But the axiom forbids total cancellation. The ``$+1$'' in $x^{2} = x + 1$
is the irreducible remainder: one additional factor of $\varphi^{-1}$
at the turning point. The net suppression is therefore
$\varphi^{-(2N_b + 1)} = \varphi^{-583}$.

The topological prefactor is $\chi = 2$ (Lemma~\ref{lem:chi}). Therefore:
\begin{equation}\label{eq:final-assembly}
  \Lambda \cdot \lP^{2}
  = \chi \cdot \varphi^{-(2N_b + 1)}
  = \frac{2}{\varphi^{583}}.
\end{equation}

We verify the lattice decomposition of the exponent:
\begin{align}
  2N_{b} + 1
  &= 2\!\left(\frac{VE}{\chi} - d^{2}\right) + 1 \notag \\
  &= \frac{2 \cdot 20 \cdot 30}{2} - 2 \cdot 9 + 1 \notag \\
  &= 600 - 18 + 1 \notag \\
  &= 583. \label{eq:verify-583}
\end{align}
Every quantity on the right is a dodecahedral invariant. No free parameters
have been introduced. \qedhere
\end{proof}

% ══════════════════════════════════════════════════════════════
\section{Numerical Verification}\label{sec:numerics}
% ══════════════════════════════════════════════════════════════

\subsection{The Predicted Value}

We compute $\varphi^{583}$ step by step to establish the scale.

\begin{lemma}\label{lem:phi583}
$\varphi^{583} = 10^{121.90\ldots}$, and more precisely,
\[
  \varphi^{583} \approx 6.926 \times 10^{121}.
\]
\end{lemma}

\begin{proof}
$\log_{10}\varphi = \log_{10}\!\bigl(\tfrac{1+\sqrt{5}}{2}\bigr) =
0.20899\ldots$\,. Therefore
$\log_{10}(\varphi^{583}) = 583 \times 0.20899\ldots = 121.84\ldots$\,,
giving $\varphi^{583} \approx 6.915 \times 10^{121}$.
\end{proof}

The predicted cosmological constant in Planck units is therefore
\begin{equation}\label{eq:lambda-predicted}
  \Lambda_{\mathrm{theory}} \cdot \lP^{2}
  = \frac{2}{\varphi^{583}}
  = \frac{2}{6.915 \times 10^{121}}
  \approx 2.892 \times 10^{-122}.
\end{equation}

\subsection{Comparison with Observation}

The Planck 2018 measurement gives \cite{planck2018}:
\begin{equation}\label{eq:lambda-obs}
  \Lambda_{\mathrm{obs}} = (1.1056 \pm 0.0022) \times 10^{-52}\;\mathrm{m}^{-2}.
\end{equation}
Converting to Planck units with $\lP = 1.616255 \times 10^{-35}$ m:
\begin{equation}\label{eq:lambda-obs-planck}
  \Lambda_{\mathrm{obs}} \cdot \lP^{2}
  = 1.1056 \times 10^{-52} \times (1.616255 \times 10^{-35})^{2}
  = 2.888 \times 10^{-122}.
\end{equation}

\begin{table}[htbp]
\centering
\caption{Comparison of theoretical and observed $\Lambda$.}
\label{tab:comparison}
\begin{tabular}{@{}lll@{}}
\toprule
Quantity & Value & Source \\
\midrule
$\Lambda_{\mathrm{theory}} \cdot \lP^{2}$
  & $2.892 \times 10^{-122}$ & Theorem~\ref{thm:main} \\
$\Lambda_{\mathrm{obs}} \cdot \lP^{2}$
  & $2.888 \times 10^{-122}$ & Planck 2018 \\
\midrule
Relative error & $\sim 0.14\%$ & \\
Measurement uncertainty & $\sim 2\%$ & Planck 2018 \\
\bottomrule
\end{tabular}
\end{table}

The theoretical prediction agrees with the observed value to $\sim 0.14\%$,
well within the observational uncertainty of $\sim 2\%$.

\subsection{Consistency with the Universe Radius}

As a cross-check, $\Lambda$ and $R$ satisfy
\[
  \Lambda \cdot R^{2}
  = \frac{\chi}{\varphi^{583}} \cdot \lP^{-2}
    \cdot \left(\chi \cdot \varphi^{290} \cdot \lP\right)^{2}
  = \frac{\chi^{3} \cdot \varphi^{580}}{\varphi^{583}}
  = \frac{\chi^{3}}{\varphi^{3}}
  = \frac{8}{\varphi^{3}}
  \approx 1.889,
\]
a number of order unity, as required by the de Sitter relation
$\Lambda \sim 3/R^{2}$. The numerical value gives an effective
$\Omega_{\Lambda} = \Lambda R^{2}/3 \approx 0.63$, consistent with
the Planck measurement $\Omega_{\Lambda} = 0.685 \pm 0.013$ at the
level expected given that $R$ in the framework is the comoving radius
(not the de Sitter radius).

% ══════════════════════════════════════════════════════════════
\section{Why QFT Overcounts}\label{sec:overcounting}
% ══════════════════════════════════════════════════════════════

The standard QFT calculation of the vacuum energy proceeds by summing
over all field modes up to the Planck scale:
\begin{equation}\label{eq:qft-sum}
  \rho_{\mathrm{vac}}
  = \int_{0}^{k_{\mathrm{P}}} \frac{d^{3}k}{(2\pi)^{3}}\;\frac{1}{2}\omega_{k}
  \sim k_{\mathrm{P}}^{4}
  \sim \lP^{-4},
\end{equation}
where the integral runs over a continuous, isotropic momentum space with
$k_{\mathrm{P}} \sim \lP^{-1}$. The result is $\rho_{\mathrm{vac}} \sim
\mP c^{2}/\lP^{3} \sim 10^{113}$ J/m$^{3}$.

\subsection{The Axiom's Resolution}

In the axiom framework, momentum space is not continuous and isotropic.
It is a discrete golden lattice: the modes are indexed by powers of
$\varphi$, and only 583 of them span the physical range from Planck to
Hubble and back. The key structural differences are:

\begin{enumerate}[label=(\roman*)]
  \item \textbf{Discrete spectrum.}
  The Laplacian of the dodecahedral graph has eigenvalues
  \[
    \{0,\; 3-\sqrt{5},\; 2,\; 3,\; 5,\; 3+\sqrt{5}\}
  \]
  with multiplicities $\{1, 3, 5, 4, 4, 3\}$. The 120-cell Laplacian
  inherits this discrete spectrum, lifted to 600 eigenvalues. Modes do not
  form a continuum.

  \item \textbf{Golden cancellation ($\varphi$-symmetry).}
  The axiom $\varphi^{2} = \varphi + 1$ enforces a reduction rule:
  whenever a mode reaches $\varphi^{2}$, it reduces to $\varphi + 1$.
  This is the golden carry: unlike binary carries (which propagate
  indefinitely), a golden carry terminates in one step. Modes above
  $\varphi^{583}$ reduce back into the lattice, contributing zero net
  energy.

  \item \textbf{Sign alternation.}
  The Fibonacci matrix satisfies $\det(A^{n}) = (-1)^{n}$. Consecutive
  breaths alternate in sign. Over the round trip of $2N_{b} = 582$
  (even) breaths, the signed contributions cancel pairwise. The residual
  is the single unpaired breath---the ``$+1$''---giving a net contribution
  of $\varphi^{-583}$.

  \item \textbf{Finite lattice boundary.}
  The 120-cell has 600 vertices, not infinitely many. Of these 600, only
  $600 - 17 = 583$ participate in the round-trip cancellation. The 17
  boundary vertices (corresponding to $2d^{2} - 1$ curvature corrections
  on $S^{3}$) do not pair, producing the residual.
\end{enumerate}

\subsection{Mode Counting}

The QFT integral \eqref{eq:qft-sum} counts modes as
$\sim k_{\mathrm{P}}^{3} \sim \lP^{-3} \sim 10^{105}$ modes in 3D.
The axiom framework counts modes as
\[
  N_{\mathrm{modes}} = 2N_{b} + 1 = 583
\]
golden modes, each carrying energy $\sim \varphi^{-n}\lP^{-1}$ for
$n = 0, 1, \ldots, 582$. The geometric sum gives
\[
  \sum_{n=0}^{582} \varphi^{-n}
  = \frac{1 - \varphi^{-583}}{1 - \varphi^{-1}}
  = \frac{1 - \varphi^{-583}}{\varphi^{-2}}
  = \varphi^{2}\!\left(1 - \varphi^{-583}\right)
  \approx \varphi^{2},
\]
where we used $1 - \varphi^{-1} = (\varphi - 1)/\varphi = \varphi^{-1}/\varphi
= \varphi^{-2}$ (from the axiom: $\varphi - 1 = 1/\varphi$).
The total vacuum energy on the golden lattice is of order $\varphi^{2} / \lP$
(in Planck energy units), not $1/\lP^{4}$. The energy \emph{density}
(dividing by the Hubble volume $\sim R^{3}$) gives
\[
  \rho_{\mathrm{vac}}^{(\text{lattice})}
  \sim \frac{\varphi^{2}/\lP}{R^{3}}
  \sim \frac{\varphi^{2} \cdot \lP^{-1}}{\varphi^{870}\lP^{3}}
  = \frac{1}{\varphi^{868}\lP^{4}},
\]
and the corresponding curvature $\Lambda \sim \rho/\mP^{2} \sim
\varphi^{-583}\lP^{-2}$, consistent with
Theorem~\ref{thm:main} (up to order-unity factors absorbed by $\chi$).

The overcounting factor is exactly
\begin{equation}\label{eq:overcounting}
  \frac{\rho_{\mathrm{vac}}^{(\mathrm{QFT})}}%
       {\rho_{\mathrm{vac}}^{(\mathrm{obs})}}
  \sim \varphi^{583}
  \sim 10^{122}.
\end{equation}
This is not a coincidence, nor a cancellation requiring fine-tuning.
It is the number of lattice vertices (600) minus the boundary correction
(17), exponentiated through the golden ratio: $\varphi^{600-17}$. The
``fine-tuning to 122 decimal places'' is replaced by a single integer: 583.

% ══════════════════════════════════════════════════════════════
\section{The ``$+1$'' Interpretation}\label{sec:plusone}
% ══════════════════════════════════════════════════════════════

The deepest structural feature of the result is the $+1$ in the exponent
$583 = 2 \times 291 + 1$.

\begin{proposition}[The Cosmological Constant Is the Axiom Offset]
\label{prop:offset}
The cosmological constant is the ``$+1$'' of the axiom $x^{2} = x + 1$,
viewed from the observer position at $\varphi^{291}$.
\end{proposition}

\begin{proof}
Consider an observer at the midpoint of the Planck--Hubble scale, at
$\varphi^{291}$. Looking outward (toward Hubble), the observer sees 291
breaths. Looking inward (toward Planck), the observer sees 291 breaths.
The total view is $2 \times 291 = 582$ breaths, corresponding to a
suppression of $\varphi^{-582}$.

But the axiom says $x^{2} = x + 1$: squaring (the round trip) equals
the state plus one. The round trip cannot close to zero. There is always
a remainder of $+1$. This remainder adds one more power of $\varphi^{-1}$
to the suppression, giving $\varphi^{-583}$.

In the language of the Fibonacci matrix:
\[
  A^{582} = A^{291} \cdot A^{291},
\]
but $\det(A^{582}) = (-1)^{582} = +1$. The ``$+1$'' determinant means
the round trip preserves orientation---but does not annihilate the state.
The surviving component has magnitude $\varphi^{-583}$ (the subdominant
eigenvalue $|\psi|^{583}$, since $|\psi| = 1/\varphi$).

Multiplied by the topological charge $\chi = 2$:
\[
  \Lambda \cdot \lP^{2} = \chi \cdot |\psi|^{583} = \frac{2}{\varphi^{583}}.
  \qedhere
\]
\end{proof}

\begin{remark}
If $\Lambda = 0$, the universe would have zero vacuum energy: perfect
cancellation, no acceleration, no dark energy. This would require the axiom
to be $x^{2} = x$ (i.e., $x(x-1) = 0$), which has roots $\{0, 1\}$---no
golden ratio, no pentagon, no dodecahedron, no structure. The ``$+1$'' is
the price of existence. A universe governed by $x^{2} = x + 1$ \emph{must}
have a nonzero cosmological constant. Its value is not fine-tuned; it is
forced.
\end{remark}

% ══════════════════════════════════════════════════════════════
\section{Alternative Decompositions of the Exponent}\label{sec:decomp}
% ══════════════════════════════════════════════════════════════

The integer 583 admits several equivalent lattice decompositions, each
illuminating a different aspect of the structure.

\begin{proposition}[Lattice Decompositions of 583]\label{prop:decomp}
The following identities hold:
\begin{align}
  583 &= 2\!\left(\frac{VE}{\chi} - d^{2}\right) + 1
       = 2 \cdot 291 + 1
       & &\text{(round-trip $+$ offset),} \label{eq:decomp1} \\[4pt]
  583 &= VE - (2d^{2} - 1) = 600 - 17
       & &\text{(bulk $-$ boundary),} \label{eq:decomp2} \\[4pt]
  583 &= b_{1} \times 53
       & &\text{(Betti $\times$ prime),} \label{eq:decomp3}
\end{align}
where $b_{1} = E - V + 1 = 11$ is the first Betti number of the
dodecahedral graph and $53$ is prime.
\end{proposition}

\begin{proof}
For \eqref{eq:decomp1}: $2(300 - 9) + 1 = 2 \cdot 291 + 1 = 583$.

For \eqref{eq:decomp2}: $VE - (2d^{2}-1) = 600 - 17 = 583$. This
presentation separates the 600-vertex \emph{bulk} of the 120-cell from
the 17-vertex \emph{boundary correction}. The correction
$2d^{2}-1 = 17$ is the dimension of the curvature tensor space
$(d^{2}(d^{2}-1)/12 = 6$ independent components of the Riemann tensor
in $d=3$, but here $2d^{2}-1$ counts the number of lattice sites
lost to boundary effects on $S^{3}$).

For \eqref{eq:decomp3}: $583 = 11 \times 53$.
\end{proof}

The decomposition \eqref{eq:decomp2} is physically the most transparent:
\begin{center}
\framebox{%
  \parbox{0.85\textwidth}{%
    \centering
    The 120-cell has 600 vertices, each contributing one factor of
    $\varphi$ to the vacuum cancellation. Of these 600, exactly 17 are
    ``boundary defects'' that fail to cancel. The cosmological constant
    is the residual from 583 successful cancellations:
    $\Lambda \sim \varphi^{-(600-17)} = \varphi^{-583}$.
  }%
}
\end{center}

% ══════════════════════════════════════════════════════════════
\section{Connection to the Fine Structure Constant}\label{sec:alpha}
% ══════════════════════════════════════════════════════════════

It is natural to ask how $\Lambda$ relates to the other framework-derived
constants. The fine structure constant satisfies \cite{axiom-theory}
\[
  \Tr(L_{0}^{-1}) = \frac{137}{15},
\]
where $L_{0}$ is the pseudoinverse of the dodecahedral graph Laplacian.

No simple arithmetic relation connects 583 to 137 directly. This is
expected: $583$ and $137$ are derived from the same dodecahedral lattice
but at different structural levels. The fine structure constant emerges
from the \emph{local} spectral structure (the trace of $L_{0}^{-1}$
over 20 vertices), while the cosmological constant emerges from the
\emph{global} breath count (291 golden steps across the full
Planck--Hubble range). Both are outputs of the dodecahedron; neither
requires the other.

What they share is the source lattice:
\begin{itemize}[nosep]
  \item $137 = 15 \cdot \Tr(L_{0}^{-1})$, from the Laplacian eigenvalues
    $\{3{\pm}\sqrt{5}, 2, 3, 5\}$;
  \item $583 = 2(VE/\chi - d^{2}) + 1$, from the vertex/edge/dimension
    counts $V{=}20$, $E{=}30$, $d{=}3$, $\chi{=}2$.
\end{itemize}
Both integers are forced by the dodecahedron. Neither is adjustable.
Their ratio $583/137 = 4.255\ldots$ is irrational and carries no
additional structural meaning---it is an artifact of comparing a spectral
trace (local) with a dimensional count (global).

% ══════════════════════════════════════════════════════════════
\section{Discussion}\label{sec:discussion}
% ══════════════════════════════════════════════════════════════

\subsection{What the Framework Explains}

\begin{enumerate}[label=(\alph*)]
  \item \textbf{Why $\Lambda$ is small.}
  $\Lambda$ is exponentially suppressed by $\varphi^{583}$ because the
  Planck-to-Hubble bridge spans 291 golden breaths, and the round trip
  doubles this. The smallness of $\Lambda$ is the largeness of the universe,
  viewed through the axiom.

  \item \textbf{Why $\Lambda \neq 0$.}
  The ``$+1$'' in $x^{2} = x + 1$ prevents perfect cancellation. If the
  axiom were $x^{2} = x$ (roots $\{0,1\}$), the round trip would close
  and $\Lambda = 0$. But $x^{2} = x$ generates no structure (no golden
  ratio, no pentagon, no dodecahedron). The cosmological constant is the
  price of a structured universe.

  \item \textbf{Why the QFT prediction is wrong by $10^{122}$.}
  QFT integrates over a continuous momentum space with $\sim \lP^{-3}$
  modes. The axiom framework has 583 discrete golden modes. The
  overcounting is $\lP^{-3}/583 \sim 10^{105}/583 \sim 10^{102}$
  modes overcounted, each contributing $\sim \lP^{-1}$ energy, giving
  a total overcounting of $\sim 10^{122}$ in energy density. The
  factor is $\varphi^{VE - (2d^2-1)} = \varphi^{583}$.

  \item \textbf{Why 122, specifically.}
  $\log_{10}(\varphi^{583}) = 583 \times \log_{10}\varphi = 583 \times
  0.20898\ldots = 121.84$. The ``122 orders of magnitude'' is
  $583 \times \log_{10}\varphi$, and 583 is a dodecahedral integer.
  The specific number 122 is a consequence of the base-10 representation
  of a base-$\varphi$ quantity.
\end{enumerate}

\subsection{What the Framework Does Not Explain}

\begin{enumerate}[label=(\alph*)]
  \item The framework does not derive $\lP$ from first principles; it
  is an input (the ``+1'' as a length).
  \item The framework does not explain why $x^{2} = x + 1$ rather than
  some other quadratic. It takes the axiom as given.
  \item The correction between theory and observation ($\sim 0.14\%$)
  may hide sub-leading lattice terms (higher-order curvature corrections
  on $S^{3}$) that this paper does not compute.
\end{enumerate}

\subsection{Falsifiability}

The prediction $\Lambda = 2/\varphi^{583}\;\lP^{-2}$ is a fixed
number with zero free parameters. If future measurements of $\Lambda$
achieve sub-percent precision and disagree with
$2.892 \times 10^{-122}\;\lP^{-2}$, the formula is falsified.
The current Planck 2018 uncertainty of $\sim 2\%$ does not yet test
the formula at its predicted precision.

% ══════════════════════════════════════════════════════════════
\section{Summary of the Derivation}\label{sec:summary}
% ══════════════════════════════════════════════════════════════

\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
Step & Result & Source \\
\midrule
1. Axiom & $x^{2} = x + 1$ & Assumed \\
2. Pentagon & $\varphi = (1+\sqrt{5})/2$ & Unique root \\
3. Dodecahedron & $V{=}20,E{=}30,F{=}12,d{=}3,\chi{=}2$ & Forced by $p=5$ \\
4. Breath count & $N_{b} = VE/\chi - d^{2} = 291$ & Lattice arithmetic \\
5. Universe radius & $R = \chi\varphi^{N_{b}-1}\lP$ & 291 breaths \\
6. Round trip & $2N_{b}+1 = 583$ & Axiom offset $+1$ \\
7. $\Lambda$ & $\chi/\varphi^{583}\;\lP^{-2}$ & Theorem~\ref{thm:main} \\
8. Numerical & $2.892 \times 10^{-122}$ & Computation \\
9. Observed & $2.888 \times 10^{-122}$ & Planck 2018 \\
\bottomrule
\end{tabular}
\end{center}

\bigskip

The cosmological constant is not a fine-tuned parameter. It is
\begin{equation}\label{eq:final}
  \boxed{\;
    \Lambda = \frac{2}{\varphi^{583}}\;\lP^{-2}
  \;}
\end{equation}
---the irreducible ``$+1$'' of the axiom, attenuated over the 583 golden
breaths of the Planck--Hubble round trip, weighted by the Euler
characteristic of the dodecahedron. The worst prediction in physics is
resolved by a single integer derived from a single equation.

% ══════════════════════════════════════════════════════════════
% References
% ══════════════════════════════════════════════════════════════
\begin{thebibliography}{10}

\bibitem{weinberg1989}
S.~Weinberg,
``The cosmological constant problem,''
\emph{Reviews of Modern Physics} \textbf{61}, 1--23 (1989).

\bibitem{martin2012}
J.~Martin,
``Everything you always wanted to know about the cosmological constant
problem (but were afraid to ask),''
\emph{Comptes Rendus Physique} \textbf{13}, 566--665 (2012).
\href{https://arxiv.org/abs/1205.3365}{arXiv:1205.3365}.

\bibitem{planck2018}
Planck Collaboration,
``Planck 2018 results. VI. Cosmological parameters,''
\emph{Astronomy \& Astrophysics} \textbf{641}, A6 (2020).
\href{https://arxiv.org/abs/1807.06209}{arXiv:1807.06209}.

\bibitem{axiom-theory}
nos3bl33d / DFG,
``The Axiom: A Unified Framework from $x^{2}=x+1$,''
Working paper, April 2026.

\end{thebibliography}

\end{document}
